---
title: "Canopy Openness and Herbaceous Ground Cover Hypo"
format: html
---

Hypothesis: As canopy openness increases, ground cover will also increase
(As canopy density increases, percent bare ground will increase)

Set things up
```{r}
rm(list = ls())
library(tidyverse)
library(here)
```


Pull in data
```{r}
gc_cd_df <- read.csv(here("Data/Ground_cover_canopy_density_summer22.csv"))
```

Workflow: 
Import -> Tidy -> Visualize/Model/Transform cycle -> Communicate 

The data I am working with have been imported and tidied in the 'Gound_cover_canopy_cleaning.qmd'

Make Some Preliminary Plots
```{r}
#all canopy variables measure the percent of canopy cover (eg 80 = 80 percent of sky is blocked by canopy = less light gets to forest floor)
#Regen_canopyapp = measure of canopy density above regen plots 
ggplot(data = gc_cd_df, aes(x = Regen_canopyapp, y = Percent_bare)) +
  geom_point()


#Regen_densiometer = measure of canopy desnity above regen plots
ggplot(data = gc_cd_df, aes(x = Regen_densiometer, y = Percent_bare)) +
  geom_point()


#Densiometer is a measure of canopy density not associated with a regen plot
ggplot(data = gc_cd_df, aes(x = Densiometer, y = Percent_bare)) +
  geom_point()


#Canopyapp is a measure of canopy density not associated with a regen plot
ggplot(data = gc_cd_df, aes(x = Canopyapp, y = Percent_bare)) +
  geom_point()

#it appears that there are similar trends between the different types of canopy measurements
```


Generate a Model 
```{r}
#first I will use a canopy measurement associated with a regen plot
mod1 <- lm(Regen_canopyapp ~ Percent_bare, data = gc_cd_df)
summary(mod1)


#and now I will make a model using a canopy measuremnet unassociated with a regen plot
mod2 <- lm(Densiometer ~ Percent_bare, data = gc_cd_df)
summary(mod2)
```

Both models produced statistically significant p-values. 

## Run correlation matrix: each row = plot, and then mean canopy density of canopyapp and mean canopy density using densiometer -> correlation should be high 



What if I found a mean canopy density (averaged from all canopy density measurements) for each plot and compared it to the mean bare ground for each plot. 
```{r}
mean_bare <- gc_cd_df |> 
  filter(Forest != "Black ash stand") |>
  group_by(Forest, Plot_num) |>
  summarise(mean_cover = mean(Percent_bare),
            mean_canopy = mean(Densiometer))

```

And then ran a test comparing the means
```{r}
mod3 <- lm(mean_canopy ~ mean_cover, data = mean_bare)
summary(mod3)
```


Make a pretty graph
```{r}
canopy_ground_cover_plot <- ggplot(data = mean_bare, aes(x = mean_canopy, y = mean_cover)) +
  geom_point(aes(color = Forest, shape = Forest), size = 2) +
  theme_bw() +
  labs(title = "Effect of Canopy Density on Ground Cover",
       x = "Mean Canopy Density",
       y = "Mean Bare Ground (%)") +
  geom_smooth(method = "lm") +
scale_color_viridis_d()

canopy_ground_cover_plot

ggsave(here("Outputs", "canopy_groundcover.png"), device = "png", height = 4, width = 6)
```





