---
title: "Deer Density and Seedling Density Hypo"
format: html
---

Pull in data
```{r}
rm(list = ls())
library(tidyverse)
library(here)

deer <- read.csv(here("Data", "deer_small_critter_abundance 1.csv"))
regen <- read.csv(here("Data", "Forest_Regeneration_17_21.csv"))
```

First, calculate regen density 
```{r}
regen_dens_plot <- regen |> 
  filter(Year == 2021) |>
  group_by(Forest, PlotNum, Quadrant) |>
  summarise(num_regen = n()) |>
  group_by(Forest, PlotNum) |>
  summarise(Mean_regen_density = sum(num_regen/3),
            SD_regen = sd(num_regen))
```

Now get deer data ready and then combine regen_dens_plot with deer data
```{r}
deer_dens <- deer |>
  filter(year == 2021) |>
  select(-c("mean_squirrels__per_100_cam_nights", "mean_small_mammals_per_100_cam_nights", "year")) |>
  mutate(forest_or_site_name = replace(forest_or_site_name, forest_or_site_name == 'Kip Tract', 'Kip tract'),
         forest_or_site_name = replace(forest_or_site_name, forest_or_site_name == 'Peavine', 'Peavine Swamp'))



#join the dataframes

deer_regen <- full_join(regen_dens_plot, deer_dens, by = c("Forest" = "forest_or_site_name", "PlotNum" = "forest_ecology_plot_number"))

```


Now I will make a preliminary plot of the relationship between variables of interest
```{r}
ggplot(data = deer_regen, aes(x = mean_deer_per_100_cam_nights, y = Mean_regen_density, color = Forest, shape = Forest)) +
  geom_point()+
  scale_color_viridis_d()
```


Now, I will run a regression on regen density and deer density
```{r}
mod1 <- lm(mean_deer_per_100_cam_nights ~ Mean_regen_density, data = deer_regen)
summary(mod1)

#p-value = 0.6881, 1 and 12 DF 
```

Take out outlier deer (60) and re run model - if different explain what I did in results, talk about in discussion 
```{r}
deer_dens <- deer |>
  filter(year == 2021) |>
  select(-c("mean_squirrels__per_100_cam_nights", "mean_small_mammals_per_100_cam_nights", "year")) |>
  mutate(forest_or_site_name = replace(forest_or_site_name, forest_or_site_name == 'Kip Tract', 'Kip tract'),
         forest_or_site_name = replace(forest_or_site_name, forest_or_site_name == 'Peavine', 'Peavine Swamp')) 

no_outlier <- deer_dens |>
  slice(-13)

```

Combine data
```{r}
deer_regen_no_outlier <- full_join(regen_dens_plot, no_outlier, by = c("Forest" = "forest_or_site_name", "PlotNum" = "forest_ecology_plot_number"))
```

Plot
```{r}
ggplot(data = deer_regen_no_outlier, aes(x = mean_deer_per_100_cam_nights, y = Mean_regen_density, color = Forest, shape = Forest)) +
  geom_point()+
  scale_color_viridis_d()
```

Regression
```{r}
mod2 <- lm(mean_deer_per_100_cam_nights ~ Mean_regen_density, data = deer_regen_no_outlier)

summary(mod2)

#still has a very large p-value
```



