---
title: "Data Cleaning"
author: "Olivia"
format: html
---

Set things up
```{r}
rm(list = ls())
library(here)
library(ggplot2)
library(dplyr)
```

Suck in the data
```{r}
canopy <- read.csv(here("Data", "Ground_cover_canopy_density_summer22.csv"), stringsAsFactors = TRUE)
```

Get a summary of the data
```{r}
summary(canopy)
```

With an initial glance at the summary, I don't see any unusual values. The percents (for density and ground cover) don't go above 100/below 0.

There are the correct number of forests and plots. Lat and Long look reasonable, no outliers. 

R is recognizing our dates as dates, and all the dates look correct. 

Now I'll take a closer look at the numberic variables. 

1. Percent bare ground
```{r}
summary(canopy$Percent_bare)
```

```{r}
ggplot(canopy, aes(Percent_bare))+
     geom_histogram()+
     labs(x = "Percent Bare Ground", y = "Number of Plots")+
     theme_bw()
```
This looks reasonable. For analysis in the future, grouping this data by plot could be interesting. 


2. Percent non_vascular
```{r}
summary(canopy$Percent_nonvascular)
```

```{r}
ggplot(canopy, aes(Percent_nonvascular))+
     geom_histogram()+
     labs(x = "Percent Nonvascular Cover", y = "Number of Plots")+
     theme_bw()
```
Looks ok, would be interesting to see which forest had the higher nonvascular cover. 


3. Percent angiosperms
```{r}
summary(canopy$Percent_angiosperms)
```

```{r}
ggplot(canopy, aes(Percent_angiosperms))+
     geom_histogram()+
     labs(x = "Percent Angiosperm Cover", y = "Number of Plots")+
     theme_bw()
```
Looks fine. 

4. Percent Ferns
```{r}
summary(canopy$Percent_ferns)
```

```{r}
ggplot(canopy, aes(Percent_ferns))+
     geom_histogram()+
     labs(x = "Percent Fern Cover", y = "Number of Plots")+
     theme_bw()
```

Ferns can be an indicator of forest health so seeing which forest had a lot of ferns could be interesting. 

5. Canopy density over regen with densiometer method
```{r}
summary(canopy$Regen_densiometer)
```
```{r}
ggplot(canopy, aes(Regen_densiometer))+
     geom_histogram()+
     labs(x = "Canopy Density Above Regen Plot", y = "Number of Measurements")+
     theme_bw()
```

6. 
```{r}
summary(canopy$Regen_canopyapp)
```

```{r}
ggplot(canopy, aes(Regen_canopyapp))+
     geom_histogram()+
     labs(x = "Canopy density over regen plot (CanopyApp)", y = "Number of Measurements")+
     theme_bw()
```

Would be interesting to see if there is a statistically significant difference between these two different methods of measuring canopy density. 

7. 
```{r}
summary(canopy$Densiometer)
```

```{r}
ggplot(canopy, aes(Densiometer))+
     geom_histogram()+
     labs(x = "Canopy density over each quadrant (densiometer)", y = "Number of Measurements")+
     theme_bw()
```
Would be interesting to group by plot/forest. 

8. 
```{r}
summary(canopy$Canopyapp)
```

```{r}
ggplot(canopy, aes(Canopyapp))+
     geom_histogram()+
     labs(x = "Canopy density over each quadrant (canopyapp)", y = "Number of Measurements")+
     theme_bw()
```

9. 
```{r}
print(canopy$Herb_spp)
```

There is nothing in this initial run through of the data that makes me suspicious. Since I didn't change anything, no need to save a new data file. 


Trying to commit

