---
title: "Final Multiple Regression "
format: html
---

## Organize Data for Regression 

Set things up
```{r}
rm(list = ls())
library(tidyverse)
library(here)

Regen <- read.csv(here("Data", "Cleaned", "regen_diversity_canopy_ground.csv"), stringsAsFactors = TRUE)

Trees <- read.csv(here("Data", "Cleaned", "tree_div_canopy_ground.csv"), stringsAsFactors = TRUE)

Test <- read.csv(here("Data", "Cleaned", "Regen_div.csv"), stringsAsFactors = TRUE)
```


Rename Simpson and Shannon and SR variables to be specific to trees and regen
```{r}
#rename regen first
Regen <- Regen |>
  rename(Simpson_regen = Simpson,
         Shannon_regen = Shannon,
         SR_regen = SR)


#trees second 
Trees <- Trees |>
  rename(Simpson_trees = Simpson,
         Shannon_trees = Shannon,
         SR_trees = SR)
```


Join the dataframes
```{r}
tree_regen_div <- full_join(Trees, Regen, by = c("Forest" = "Forest", "Plot_num" = "Plot_num", "Mean_ground_cover" = "Mean_ground_cover", "canopy_density" = "canopy_density"))

write.csv(tree_regen_div, here("Data", "Cleaned", "tree_regen_div"), row.names = FALSE)
```

Now, pull in soil data
```{r}
Soil <- read.csv(here("Data", "Cleaned", "EnviroVars2021.csv"))

ggplot(data = Soil, aes(x = mCEC)) +
  geom_histogram()

ggplot(data = Soil, aes(x = m_pH)) +
  geom_histogram()

ggplot(data = Soil, aes(x = mK)) +
  geom_histogram()
#won't use

ggplot(data = Soil, aes(x = mMg)) +
  geom_histogram()
#won't use

ggplot(data = Soil, aes(x = mNitrate)) +
  geom_histogram()

ggplot(data = Soil, aes(x = mSOM)) +
  geom_histogram()

#I will look at mCEC, m_pH, mNitrate, and mSOM
soil_info <- select(Soil, c("Forest", "Plot", "mCEC", "m_pH", "mNitrate", "mSOM"))

#rename kip tract to just kip 
soil_info <- soil_info |>
  mutate(Forest = replace(Forest, Forest == 'Kip tract', 'Kip'))
```

Add soil to dataframe
```{r}
div_soil <- full_join(tree_regen_div, soil_info, by = c("Forest" = "Forest", "Plot_num" = "Plot"))
```

Now, calculate tree basal area 
```{r}
Basal <- read.csv(here("Data", "Forest_Trees_and_Regen_2021.csv"))

Find_basal <- Basal |>
  filter(Type == "Tree") |>
  mutate(DBH_m = DBH_cm/100,
         radius_m = DBH_m/2,
         area_m2 = pi*(radius_m*radius_m)) |>
  group_by(Forest, Plot) |>
  summarise(total_basal_m2 = sum(area_m2),
            median_basal_m2 = median(area_m2),
            sd_basal_m2 = sd(area_m2)/sqrt(sum(!is.na(area_m2))))

#find basal averaged per forest
basal_by_forest <- Find_basal |>
  group_by(Forest) |>
  summarise(Mean_total_basal = mean(total_basal_m2),
            Mean_median_basal = mean(median_basal_m2),
            Sd_total_basal = sd(total_basal_m2),
            Sd_median_basal = sd(median_basal_m2))
```


Add basal area to dataframe
```{r}
#first rename kip
Find_basal <- Find_basal |>
  mutate(Forest = replace(Forest, Forest == 'Kip tract', 'Kip'))


div_soil_basal <- full_join(div_soil, Find_basal, by = c("Forest" = "Forest", "Plot_num" = "Plot"))
```

Now, pull in mammal data for mammal diversity 
```{r}
mammal_div <- read.csv(here("Data", "mammal_data_for_vegan.csv"))
```

Use vegan to calculate diveristy 
```{r}
library(vegan)

shannon <- diversity(mammal_div[,3:25], index = "shannon")

Mammal_Diversity <- data.frame(Forest = character(15),
                         Plot = numeric(15),
               Shannon_mammals = numeric(15))

Mammal_Diversity$Forest <- mammal_div$forest_or_site_name
Mammal_Diversity$Plot <- mammal_div$forest_ecology_plot_number
Mammal_Diversity$Shannon_mammals <- shannon
```


Now add shannon for mammals to dataframe
```{r}
Mammal_Diversity <- Mammal_Diversity |>
  mutate(Forest = replace(Forest, Forest == 'Kip Tract', 'Kip'))

Mammal_Diversity <- Mammal_Diversity |>
  mutate(Forest = replace(Forest, Forest == 'Peavine', 'Peavine Swamp'))

div_soil_basal_mammal <- full_join(div_soil_basal, Mammal_Diversity, by = c("Forest" = "Forest", "Plot_num" = "Plot"))
```


Pull in deer/squirrel/small mammal density
```{r}
animals <- read.csv(here("Data", "deer_small_critter_abundance 1.csv"))
```

Mean density
```{r}
density <- animals |>
  group_by(forest_or_site_name, forest_ecology_plot_number) |>
  summarize(mean_deer_density_per100 = mean(mean_deer_per_100_cam_nights, na.rm = T),
            mean_squirrel_density_per100 = mean(mean_squirrels__per_100_cam_nights, na.rm = T),
            mean_small_mammal_per100 = mean(mean_small_mammals_per_100_cam_nights, na.rm = T))

#for methods, the animal density measurements include camera data from 2021-2023. 
```

Add to big dataframe
```{r}
#rename forest and plot variables 
density <- density |>
  rename(Forest = forest_or_site_name,
         Plot_num = forest_ecology_plot_number)


#rename kip and peavine 
density <- density |>
  mutate(Forest = replace(Forest, Forest == 'Kip Tract', 'Kip'))

density <- density |>
  mutate(Forest = replace(Forest, Forest == 'Peavine', 'Peavine Swamp'))

full_for_regression <- full_join(div_soil_basal_mammal, density, by = c("Forest" = "Forest", "Plot_num" = "Plot_num"))

write.csv(full_for_regression, here("Data", "Cleaned", "data_for_table.csv"), row.names = FALSE)
```

Delete columns I don't need
```{r}
correlation_df <- full_for_regression |>
  select(-c("Simpson_regen", "Simpson_trees", "sd_basal_m2", "SR_trees"))

write.csv(correlation_df, here("Data", "Cleaned", "data_for_regression.csv"), row.names = FALSE)
```

## Run the Regression

Notes from meeting with Dr. Lock: 

Run cor matrix on full dataset - response w forest, and response w mammal 

If cor is > 0.7, pick one of the variables 

How should I model? Best subsets or stepwise (nvmax, num predictor, at 5). 
Do stepwise after best subsets 

Generally, select best model using aic? How should I generate those values?  
- get aic with stepwise output 

Limitation with degrees freedom - should I do one set of forest models, one set of mammal models? 

Assumptions - are data normal? - (just check for major outliers, distributions of shannon and predictor variables w scatterplots) 

Descriptive project, not formal hypo testing 

Using shannon of trees for forest health proxy - could use something else, like total forest basal area
- see if basal is highly related to shannon. if it's different, than might be measuring something different and could have different predictors in the model 

lm

3 best forest measures - does the model get better by throwing in a mammal measurement? 

pick forest based model, look at residuals and save as a new variable, see if there is a mammal based predictor that is strongly correlated to residuals 

could also do best subsets with just mammal based 

Clean things up and pull in data
```{r}
rm(list = ls())
library(here)
library(tidyverse)
library(leaps)
source(here("ShowSubsets.R"))

DF <- read.csv(here("Data", "Cleaned", "data_for_regression.csv"))
```

Run a full correlation matrix
```{r}
for_matrix <- DF |> select(-c(Forest, Plot_num))
full_matrix <- cor(for_matrix)
round(full_matrix, 3)

write.csv(full_matrix, here("Outputs", "Cor_matrix.csv"), row.names = FALSE)


#can see that SR_regen has a correlation with mSOM of -0.778
#mCEC has a 0.849 correlation with mSOM and a -0.839 with m_pH
#m_pH and mSOM have a -0.937
#mSOM Shannon_regen have -0.835

#to avoid correlation between predictors, seems like it would be a good idea to choose just one of the soil measurements. Since mSOM has high correlation with SR_regen and Shannon_regen, maybe I won't use that one, and instead will use either mCEC or m_pH. 
#of the mammal variables, seems like squirrels could be best predictor for shannon of trees. Mean_tree_density and median_basal_m2, as well as Shannon of mammals look pretty good. 

#chose to dump mSOM, SR_regen, m_pH 
```


Run a cor matrix just on forest predictors


Run a cor matrix just on mammal predictors

Run a best subsets on full dataset to see what happens, using max number of predictors = 5. 
```{r}
subset_full <- regsubsets(Shannon_trees ~  Mean_ground_cover + canopy_density + Mean_regen_density + Mean_tree_density + Shannon_regen + mCEC + mNitrate + total_basal_m2 + median_basal_m2 + Shannon_mammals + mean_deer_density_per100  + mCEC + mean_squirrel_density_per100 + mean_small_mammal_per100, data = DF, method = "forward")

ShowSubsets(subset_full)
```

```{r}
subset_5 <- regsubsets(Shannon_trees ~  .,nbest = 5, data = DF, method = "forward")

ShowSubsets(subset_5)


subset_5 <- regsubsets(Shannon_trees ~  Mean_ground_cover + canopy_density + Mean_regen_density + Mean_tree_density + Shannon_regen + mCEC  + mNitrate + total_basal_m2 + median_basal_m2 + Shannon_mammals + mean_deer_density_per100  + mean_squirrel_density_per100 + mean_small_mammal_per100, data = DF, method = "forward", nvmax = 5)

ShowSubsets(subset_5)
```

Run a best subsets just with tree data
```{r}
subset_tree <- regsubsets(Shannon_trees ~  Mean_ground_cover + canopy_density + Mean_regen_density + Mean_tree_density + Shannon_regen + mNitrate + mSOM + total_basal_m2 + median_basal_m2, data = DF, method = "forward", nvmax = 5)

ShowSubsets(subset_tree)

#this shows that the model with the lowest Cp (0.26) includes the predictors mNitrate, Shannon_regen, and Mean_tree_density

mod_forest <- lm(Shannon_trees ~ mNitrate + Shannon_regen + Mean_tree_density, data = DF)
summary(mod_forest)
#p-value: 0.008774, Adjusted R-squared:  0.5394 
```

Run a best subsets with just mammal predictors
```{r}
subset_mammal <- regsubsets(Shannon_trees ~ Shannon_mammals + mean_deer_density_per100  + mean_squirrel_density_per100 + mean_small_mammal_per100, data = DF, method = "forward", nvmax = 5)

ShowSubsets(subset_mammal)
#the model with the best Cp (4.65) includes the predictors mean_small_mammal_per100, mean_squirrel_density_per100, and Shannon_mammals. The single best predictor on it's own is mean_squirrel_density_per100. 
```


Run a best subsets with the best from each of the separate ones
```{r}
subset_combine <- regsubsets(Shannon_trees ~ Shannon_mammals  + mean_squirrel_density_per100 + mean_small_mammal_per100 + mNitrate + Shannon_regen + Mean_tree_density, data = DF, method = "forward", nvmax = 5)

ShowSubsets(subset_combine)

#Cps are not as good
```

So, going back to the plain forest model, I will see what happens if I add a mammal predictor to that model
```{r}
mod_forest_mammal <- lm(Shannon_trees ~ mNitrate + Shannon_regen + Mean_tree_density + Shannon_mammals, data = DF)

summary(mod_forest_mammal)

#with Shannon_mammals added, the p-value: 0.006754, Adjusted R-squared:  0.621 
```

Run a stepwise regression on this combination of predictors to get AIC
```{r}
mod_forest_mammal_step <- lm(Shannon_trees ~ mNitrate + Shannon_regen + Mean_tree_density + Shannon_mammals, data = DF)

step(mod_forest_mammal_step)

#AIC = -39.64
#I also notice that Shannon_mammals has a negative coefficient 


#run AIC table on all best model options - AIC of less than 2 away from lowest is considered equivalent 
```


Run another set of models with Regen Density as response variable (much more variation)

```{r}
subset_regen <- regsubsets(Mean_regen_density ~  Shannon_trees + Mean_ground_cover + canopy_density + Mean_tree_density + Shannon_regen + mCEC + mNitrate + total_basal_m2 + median_basal_m2 + Shannon_mammals + mean_deer_density_per100  + mCEC + mean_squirrel_density_per100 + mean_small_mammal_per100, data = DF, method = "forward", nvmax = 5)

ShowSubsets(subset_regen)
```


The best three models according to the regsubsets and the Cp output
```{r}
mod_forest_regen1 <- lm(Mean_regen_density ~ mCEC + mNitrate + median_basal_m2 + Shannon_mammals + Mean_ground_cover, data = DF)

summary(mod_forest_regen1)

mod_forest_regen2 <- lm(Mean_regen_density ~ mCEC + mNitrate + median_basal_m2 + Shannon_mammals, data = DF)

summary(mod_forest_regen2)

mod_forest_regen3 <- lm(Mean_regen_density ~ mCEC + mNitrate + median_basal_m2, data = DF)

summary(mod_forest_regen3)
```

Create AIC table
```{r}
library(AICcmodavg)

models <- list(mod_forest_regen1, mod_forest_regen2, mod_forest_regen3)

mod.names <- c('model1', 'model2', 'model3')

aictab(cand.set = models, modnames = mod.names)

#getting errors here because of package issues
```


